Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-06-06 doc/invt.tex
small changes

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the
resulting distribution, $\pi_i(t)$,
%
the real deviation, $x_i(t)$, of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014},
%
\begin{equation}
  x_i(t) \equiv v_i(t) - \bav{v}(t)
  %=
  %v_i - \sum_{j = 1}^n \rho_j \, v_j
  ,
\label{eq:x_def}
\end{equation}
%
where
%
\begin{equation}
  \bav{v}(t)
  = \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vbar_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$
(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
%
The total error is given by
%
\begin{equation}
  \Err(t)
  =
  \sum_{i = 1}^n \rho_i \, \left\langle x_i^2(t) \right\rangle
  .
\label{eq:error_def}
\end{equation}

%Although the above error weighting distribution, $\pmb\rho$,
%is usually flat,
%i.e. $\rho_i \equiv 1/n$,
%in practice,
%a non-flat distribution
%is sometimes more convenient in analyzing
%some multiple-bin updating schemes later on.
%
Note that the distribution, $\pmb\rho$, is arbitrary
and independent of
the target distribution,
$\mathbf p = (p_1, \dots, p_n)$,
although it is usually flat.\footnote{In
well-tempered metadynamics,
$\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
  barducci2008, dama2014}:
$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
whereas $\pmb\rho$ is flat.}
%
From Eqs. \eqref{eq:x_def} and \eqref{eq:vbar_def},
we have
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, x_i(t) = 0
  .
\label{eq:rhox_sum}
\end{equation}
%

\subsubsection{Differential equation}

To proceed, we will
approximate the finite difference Eq. \eqref{eq:wl_update}
by a differential equation\footnote{This
approximation requires us to shift the origin of time by $1$ in some cases.}
%
\begin{equation}
  \dot u_i(t)
  \equiv
  \frac{ d u_i(t) } { dt }
  =
  \alpha(t) \, \frac{ h_i(t) } { p_i },
  \label{eq:vt_diffeq}
\end{equation}
%
with
%
$
h_i(t) = \delta_{i, i(t)}
$
%\begin{equation}
%  h_i(t) = \delta_{i, i(t)}
%  ,
%  \label{eq:h_def}
%\end{equation}
%
being the instantaneous histogram,
which is equal to $1.0$
for the bin $i(t)$ at time $t$
or zero otherwise.
%
The shifted bias potential, $v_i(t)$,
follows the same differential equation
as the shift is a constant.

We can split $h_i(t)$ into a deterministic averaged part
and a random fluctuation part, $\zeta_i(t)$,
%
\begin{equation}
  h_i(t) = \langle h_i(t) \rangle + \zeta_i(t).
  \label{eq:h_split}
\end{equation}
%
The deterministic part can be related
to an ``ensemble average'',
where the ensemble consists of many similar simulation copies
sharing the same schedule $\alpha(t)$
and the same bias potential at time $t$.
%
The initial states and the stochastic forces
during the process may differ, however.

For sufficiently small $\alpha(t)$,
the bias potential remains roughly the same for a short period.
%
Then,
the sampling process is approximately Markovian, % of a finite order,
and we may assume Eq. \eqref{eq:pi_p_V}
for the deterministic part
%
\begin{align}
  \langle h_i(t) \rangle
  &\approx
  \pi_i(t)
  =
  \frac{                p_i \, e^{-x_i(t)} }
       { \sum_{j = 1}^n p_j \, e^{-x_j(t)} }
  \notag
  \\
  &\approx
  p_i \,
  \left[
    1 - x_i(t) + \sum_{j = 1}^n p_j \, x_j(t)
  \right]
  ,
  \label{eq:h_ave}
\end{align}
%
where we have assumed the smallness
of $x_i(t)$ in the linear expansion
and used $\sum_{j=1}^n p_j = 1$
on the second line.
%
\note{
The second step follows from
$$
\frac{ \langle h_i(t) \rangle }
     { p_i }
\approx
\frac{                       1 - x_i  }
     { \sum_{ r = 1 }^n p_j (1 - x_j) }
=
\frac{                       1 - x_i  }
     { 1 - \sum_{ j = 1 }^n p_j \, x_j }
.
$$
}%

%
For the histogram fluctuation part,
the average is zero by definition,
$\langle \zeta_i(t) \rangle = 0$,
%
and we will approximate it as white noise, i.e.
%The noise is not necessarily white,
%and the autocorrelation function
%depend only on the time difference
%
\begin{equation}
  \left\langle
    \zeta_i(t) \, \zeta_j(t')
  \right\rangle
  =
  %\sigma_{ij}(t - t')
  G^{(0)}_{ij} \, \delta(t - t')
  ,
  \notag
  %\label{eq:zeta_zeta_correlation}
\end{equation}
%
for some value of $G^{(0)}_{ij}$,
and $\delta(t)$ is the Dirac's delta function.
%where $\sigma_{ij}(t)$ is an even function
%that vanishes at large $t$.
%
This approximation is valid if the simulation period, $T$,
is much longer than the autocorrelation time of noise.
%
For non-white noise,
the equivalent value of $G^{(0)}_i$ can be approximately obtained as
$\sum_{t = -\infty}^\infty \langle \zeta_i(t) \, \zeta_j(0) \rangle$.
%
\note{More explicitly,
if the transition matrix of a period, $\tau$,
is $T^\tau_{ij}$,
then
$
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  \langle h_i(t) \, h_j(t') \rangle
  -
  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
  =
  T^{t - t'}_{ij} \, \pi_j - \pi_i \, \pi_j.
$
}%

%The average growth rate of $v(t)$ is given by
%\begin{equation}
%  \frac{ d \bav{v} (t) }
%       { d t }
%  =
%  \alpha(t)
%  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
%  .
%\label{eq:dvbardt}
%\end{equation}
%

From Eqs.
\eqref{eq:v_def}-\eqref{eq:vbar_def},
\eqref{eq:rhox_sum}-\eqref{eq:h_ave},
%\eqref{eq:v_def},
%\eqref{eq:x_def},
%\eqref{eq:vbar_def},
%\eqref{eq:rhox_sum},
%\eqref{eq:vt_diffeq},
%\eqref{eq:h_split}
%\eqref{eq:h_ave},
%\eqref{eq:dvbardt},
we get % a set of decoupled equations
%
\begin{equation}
  \dot x_i(t)
  =
  -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
  ,
\label{eq:dxdt_singlebin}
\end{equation}
%
where the shifted histogram fluctuation
%
\begin{equation}
  \xi_i(t)
  \equiv
  \frac{ \zeta_i(t) }
       { p_i }
  -
  \sum_{ j = 1 }^n
    \rho_j
    \frac{ \zeta_j(t) }
         {     p_j    }
  ,
\label{eq:xi_def}
\end{equation}
%
is also governed by white noise as is $\zeta_i(t)$:
%i.e. we may write
\begin{equation}
  \langle \xi_i(t) \, \xi_i(t') \rangle
  = G_i \, \delta(t - t')
  ,
  \label{eq:G_def}
\end{equation}
for some $G_i$.
%
\note{The derivation:
$$
\begin{aligned}
  \dot x_i
  &\stackrel{\mathrm{Eq.\; \eqref{eq:x_def}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \dot v_i - \bav{\dot v}
  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:vbar_def}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left(
    \frac{ h_i } { p_i }
    -
    \sum_{ j = 1 }^n
      \rho_j \frac{ h_j } { p_j }
  \right)
  \\
  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_split} \; and \; \eqref{eq:xi_def}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left(
    \frac{ \langle h_i \rangle } { p_i }
    -
    \sum_{ j = 1 }^n
      \rho_j \frac{ \langle h_j \rangle } { p_j }
    +
    \xi_i(t)
  \right)
  \\
  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_ave} \;and\; \eqref{eq:rhox_sum}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left[ -x_i(t) + \xi_i(t) \right].
\end{aligned}
$$
}

The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
%
\begin{equation}
x_i(T)
=
x_i(0) \, e^{-q(T)}
+
\int_0^T
  \dot y\bigl( q(t) \bigr) \, \xi_i(t) \, dt,
\notag
%\label{eq:xt_solution}
\end{equation}
%
where
%
$
q(t) \equiv \int_0^t \alpha(t') \, dt',
$
%
and
%
\begin{align}
y(q')
&\equiv
e^{q' - q(T)}.
\label{eq:y_def}
\end{align}

@@

%To simplify the ensuing discussion,
Second, we limit ourselves to matrices % $\mathbf w$'s
that satisfy
the detailed balance condition,
%
\begin{equation}
  \rho_i \, w_{ij} = \rho_j \, w_{ji}
  ,
  \label{eq:w_detailedbalance}
\end{equation}
%
for the error weighting distribution, $\pmb \rho$,
introduced in Eq. \eqref{eq:vbar_def}.
%
This requires the updating matrix, $\mathbf w$
to be symmetric
for a uniform $\pmb\rho$,
or more generally,
the scaled matrix, $\sqrt{ \rho_i/\rho_j } \, w_{ij}$,
to be symmetric.
%
This symmetry allows the diagonalization
of $\mathbf w$ with a set of
eigenvectors, $\phi_{ki}$,
%
\begin{equation}
  \sum_{i = 1}^n \phi_{ki} \, w_{ij}
  =
  \lambda_k \, \phi_{kj}
  ,
\label{eq:eig_w}
\end{equation}
%
satisfying the orthonormal conditions\cite{vankampen}:
%
\begin{align}
  \sum_{k = 0}^{n - 1}
    \phi_{ki} \, \phi_{kj}^*
  &=
  \delta_{ij} \, \rho_i,
  \label{eq:eig_orthonormal_cols}
  \\
  \sum_{i = 1}^n
    \frac{ \phi_{ki} \, \phi_{li}^* }
         { \rho_i }
  &=
  \delta_{kl}
  ,
  \label{eq:eig_orthonormal_rows}
\end{align}
%
where the superscript $^*$ denotes the complex conjugate,
and we have set the index of the first eigenmode to $0$
instead of $1$ for later convenience.

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \label{eq:w_balance}
\end{equation}
%
We will label this eigenvector as the first one,
and
%
\begin{equation}
  \phi_{0i} = \rho_i,
\label{eq:eigenmode0}
\end{equation}
%
with $\lambda_0 = 1$.
%
Clearly, Eq. \eqref{eq:eigenmode0}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 0$,
while the orthogonality demands
%
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \delta_{k0}
  .
\label{eq:ortho1}
\end{equation}
%
We may write the updating matrix in $\phi_{ki}$'s
as\cite{bussi2006}
%
\begin{equation}
  w_{ij}
  =
  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
  \lambda_k \, \phi_{ki}^* \, \phi_{kj}
  .
  \notag
\end{equation}

@@

We can further define
a generalized Fourier transform, $\mathcal{F}$,
for variable, $A_i$,
%
\begin{equation}
  {\tilde A}_k
  \equiv \mathcal{F}[A_i]_k
  \equiv \sum_{i = 1}^n \phi_{ki} \, A_i
  ,
  \label{eq:gft_def}
\end{equation}
and the inverse transform is
%
\begin{equation}
  A_i =
  \frac{1}{\rho_i} \sum_{k = 1}^n \phi^*_{ki} \, \tilde{A}_k
  .
  \label{eq:gft_inv}
\end{equation}
%
The multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
can be diagonalized as
%
$$
{\tilde v}_k(t + 1) =
{\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
\mathcal{F}\bigl[ h_i(t) / p_i \bigr]_k
,
$$
or in the continuous-time form,
%where
%%
%$$
%\eta_i \equiv \frac{ h_i(t) } { p_i } - \sum_{j = 1}^n \rho_j \frac{ h_j(t) }{ p_j }.
%$$
%
\begin{equation}
  \dot{\tilde v}_k(t)
  =
  \alpha(t) \, \lambda_k \, \mathcal{F} \bigl[ h_j(t)/ p_i  \bigr]_k
  .
  \label{eq:ut_diffeq_mbin}
\end{equation}

In particular, for the first eigenmode,
we have $\tilde{v}_0 \equiv v_\mathrm{av}$,
and
%
\begin{align*}
  \bav{\dot v}(t)
  =
  \alpha(t)
  \sum_{r = 1}^n \rho_r
    \, \frac{ h_r(t) } { p_r }
  =
  \alpha(t)
  \sum_{j = 1}^n w_{ij}
  \sum_{r = 1}^n \rho_r
    \, \frac{ h_r(t) } { p_r }
  ,
\end{align*}
%
from Eq. \eqref{eq:eigenmode0},
%
%
and by Eqs. \eqref{eq:x_def}, \eqref{eq:rhox_sum} and \eqref{eq:h_ave},
%
$$
\begin{aligned}
  \dot x_i(t)
  &= \alpha(t) \sum_{j=1}^n w_{ij}
  \left(
    \frac{ h_j(t) } { p_j }
    -
    \sum_{r = 1}^n
      \rho_r \frac{ h_r(t) } { p_r }
  \right)
  \\
  &\approx
  -\alpha(t) \sum_{j = 1}^n
  w_{ij} \bigl[ x_j(t) - \xi_j (t) \bigr]
  .
\end{aligned}
$$
%where
%we have expanded the right-hand side
%in linear terms of $x_j$
%on the second line.

Diagonalizing the matrix, $\mathbf w$,
using the eigenvectors defined in Eqs.
\eqref{eq:eig_w}-\eqref{eq:eig_orthonormal_rows}
yields a set of decoupled equations
%
\begin{equation}
  \dot{\tilde x}_k(t)
  =
  -\alpha(t) \, \lambda_k \,
  \left[{\tilde x}_k(t) - {\tilde \xi}_k(t) \right]
  .
  \notag
  %\label{eq:yt_diffeq}
\end{equation}
%
%are the deviation and noise (histogram fluctuation)
%of mode $k$, respectively.

@@

The error defined in Eq. \eqref{eq:error_def}
can be rewritten in terms of the eigenmodes of $\mathbf w$,
%
\begin{align}
  \Err
  \equiv
  \sum_{i = 1}^n \rho_i \,
    \left\langle x_i^2 \right\rangle
  =
  \sum_{i, j = 1}^n \sum_{k=0}^{n-1}
   \phi_{ki} \, \phi_{kj}^* \,
    \left\langle x_i \, x_j \right\rangle
  =
  \sum_{k = 0}^{n - 1}
    \left\langle
      \bigl| {\tilde x}_k^2 \bigr|
    \right\rangle
  ,
\notag
\end{align}
%
where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
%
From Eq. \eqref{eq:eigenmode0},
the first eigenmode,
${\tilde x}_0 = \sum_{j=1}^n \rho_j \, x_j$,
vanishes identically by Eq. \eqref{eq:rhox_sum}.
%
Thus, we can start the last sum %in Eq. \eqref{eq:xt2_sum}
from $k = 1$,
%
\begin{equation}
  \Err(T)
  =
  \sum_{k = 1}^{n - 1}
    \left\langle
      \bigl| {\tilde x}_k^2(T) \bigr|
    \right\rangle
  .
\label{eq:xt2_sum}
\end{equation}

>>>>

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014}.
%
We define a bin-average-deducted bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  ,
\label{eq:x_def}
\end{equation}
%
where the bin average, $\bav{v}(t)$, is defined as
%
\begin{equation}
  \bav{v}(t)
  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vav_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$
(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
%
The total error is given by
%
\begin{equation}
  \Err(t)
  =
  \sum_{i = 1}^n \rho_i \, \left\langle x_i^2(t) \right\rangle
  .
\label{eq:error_def}
\end{equation}

%Although the above error weighting distribution, $\pmb\rho$,
%is usually flat,
%i.e. $\rho_i \equiv 1/n$,
%in practice,
%a non-flat distribution
%is sometimes more convenient in analyzing
%some multiple-bin updating schemes later on.
%
Note that the distribution, $\pmb\rho$, is arbitrary
and independent of
the target distribution,
$\mathbf p = (p_1, \dots, p_n)$,
although it is usually flat.\footnote{In
well-tempered metadynamics,
$\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
  barducci2008, dama2014}:
$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
whereas $\pmb\rho$ is flat.}
%
From Eqs. \eqref{eq:x_def} and \eqref{eq:vav_def},
we have
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, x_i(t) = 0
  .
\label{eq:rhox_sum}
\end{equation}
%

\subsubsection{\label{sec:sbin_diffeq}
Differential equation}

To proceed, we will
approximate the finite difference Eq. \eqref{eq:wl_update}
by a differential equation\footnote{This
approximation requires us to shift the origin of time by $1$ in some cases.}
%
\begin{equation}
  \dot u_i(t)
  \equiv
  \frac{ d u_i(t) } { dt }
  =
  \alpha(t) \, \frac{ h_i(t) } { p_i }
  \equiv
  \alpha(t) \, \eta_i(t)
  ,
  \label{eq:vt_diffeq}
\end{equation}
%
with
%
$
h_i(t) = \delta_{i, i(t)}
$
%\begin{equation}
%  h_i(t) = \delta_{i, i(t)}
%  ,
%  \label{eq:h_def}
%\end{equation}
%
being the instantaneous histogram,
which is equal to $1.0$
for the bin $i(t)$ at time $t$
or zero otherwise,
and we have defined a relative histogram
$\eta_i(t) \equiv h_i(t) /p_i$
in the last step.
%
The shifted bias potential, $v_i(t)$,
follows the same differential equation
as the shift is a constant.

We now try to split the histogram into a deterministic averaged part
and a random fluctuation part, $\zeta_i(t)$,
%
\begin{equation}
  \eta_i(t) = \langle \eta_i(t) \rangle + \zeta_i(t).
  \label{eq:h_split}
\end{equation}
%
The deterministic part can be related
to an ``ensemble average'',
where the ensemble consists of many similar simulation copies
sharing the same schedule $\alpha(t)$
and the same bias potential at time $t$.
%
The initial states and the stochastic forces
during the process may differ, however.

For sufficiently small $\alpha(t)$,
the bias potential remains roughly the same for a short period.
%
Then,
the sampling process is approximately Markovian, % of a finite order,
and we may assume Eq. \eqref{eq:pi_p_V}
for the deterministic part
%
\begin{align}
  \langle \eta_i(t) \rangle
  %&=
  %\frac{ \langle h_i(t) } { p_i } \rangle
  &\approx
  \frac{ \pi_i(t) } { p_i }
  =
  \frac{                       e^{-x_i(t)} }
       { \sum_{j = 1}^n p_j \, e^{-x_j(t)} }
  \notag
  \\
  &\approx
  1 - x_i(t) + \sum_{j = 1}^n p_j \, x_j(t)
  ,
  \notag
  %\label{eq:h_ave}
\end{align}
%
where we have assumed the smallness
of $x_i(t)$ in the linear expansion
and used $\sum_{j=1}^n p_j = 1$
on the second line.
%
\note{
The second step follows from
$$
\frac{ \langle h_i(t) \rangle }
     { p_i }
\approx
\frac{                       1 - x_i  }
     { \sum_{ r = 1 }^n p_j (1 - x_j) }
=
\frac{                       1 - x_i  }
     { 1 - \sum_{ j = 1 }^n p_j \, x_j }
.
$$
}%
%
Deducting the bin average [similar to Eq. \eqref{eq:vav_def}],
we have
%
\begin{align}
  \langle \eta_i(t)  - \eta_\mathrm{av}(t) \rangle
  \equiv
  \langle \eta_i(t) \rangle -
  \sum_{j=1}^n \rho_j \langle \eta_j(t) \rangle
  &\approx
  - x_i(t)
  .
  \label{eq:sh_ave}
\end{align}

%
For the histogram fluctuation part,
the average is zero by definition,
$\langle \zeta_i(t) \rangle = 0$,
%
and we will approximate it as white noise.
%i.e.
%%The noise is not necessarily white,
%%and the autocorrelation function
%%depend only on the time difference
%%
%\begin{equation}
%  \left\langle
%    \zeta_i(t) \, \zeta_j(t')
%  \right\rangle
%  =
%  %\sigma_{ij}(t - t')
%  G^{(0)}_{ij} \, \delta(t - t')
%  ,
%  \notag
%  %\label{eq:zeta_zeta_correlation}
%\end{equation}
%%
%for some value of $G^{(0)}_{ij}$,
%where $\sigma_{ij}(t)$ is an even function
%that vanishes at large $t$.
%
This approximation is valid if the simulation period, $T$,
is much longer than the autocorrelation time of noise.
%
The bin-average-deducted fluctuation
%
\begin{equation}
  \xi_i(t)
  \equiv
  \zeta_i(t) - \zeta_\mathrm{av}(t)
  \equiv
  \zeta_i(t)
  -
  \sum_{ j = 1 }^n
    \rho_j \, \zeta_j(t)
  ,
\label{eq:xi_def}
\end{equation}
%
would also be governed by white noise as is $\zeta_i(t)$:
%i.e. we may write
\begin{equation}
  \langle \xi_i(t) \, \xi_i(t') \rangle
  = G_i \, \delta(t - t')
  ,
  \label{eq:G_def}
\end{equation}
for some $G_i$,
and $\delta(t)$ is the Dirac's delta function.
%
For non-white noise,
the equivalent value of $G_i$ can be approximately obtained as
$\sum_{t = -\infty}^\infty \langle \xi_i(t) \, \xi_i(0) \rangle$.
%
\note{More explicitly,
if the transition matrix of a period, $\tau$,
is $T^\tau_{ij}$,
then
$
  p_i p_j \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  \langle h_i(t) \, h_j(t') \rangle
  -
  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
  =
  T^{t - t'}_{ij} \, \pi_j - \pi_i \, \pi_j.
$
}%

From Eqs.
\eqref{eq:x_def}, \eqref{eq:vt_diffeq}-\eqref{eq:xi_def},
%\eqref{eq:v_def},
%\eqref{eq:x_def},
%\eqref{eq:vav_def},
%\eqref{eq:rhox_sum},
%\eqref{eq:vt_diffeq},
%\eqref{eq:h_split}
%\eqref{eq:h_ave},
%\eqref{eq:dvbardt},
we get % a set of decoupled equations
%
\begin{equation}
  \dot x_i(t)
  =
  \alpha(t) \, \left[ \eta_i(t) - \eta_\mathrm{av} \right]
  \approx
  -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
  ,
  \notag
  %\label{eq:dxdt_singlebin}
\end{equation}
%
and the formal solution is % of Eq. \eqref{eq:dxdt_singlebin} is
%
\begin{equation}
x_i(T)
=
x_i(0) \, e^{-q(T)}
+
\int_0^T
  \dot y\bigl( q(t) \bigr) \, \xi_i(t) \, dt,
\notag
%\label{eq:xt_solution}
\end{equation}
%
where
%
$
q(t) \equiv \int_0^t \alpha(t') \, dt',
$
%
and
%
\begin{align}
y(q')
&\equiv
e^{q' - q(T)}.
\label{eq:y_def}
\end{align}

@@

%To simplify the ensuing discussion,
Second, we limit ourselves to matrices % $\mathbf w$'s
that satisfy
the detailed balance condition,
%
\begin{equation}
  \rho_i \, w_{ij} = \rho_j \, w_{ji}
  ,
  \label{eq:w_detailedbalance}
\end{equation}
%
for the error weighting distribution, $\pmb \rho$,
introduced in Eq. \eqref{eq:vav_def}.
%
This requires the updating matrix, $\mathbf w$
to be symmetric
for a uniform $\pmb\rho$,
or more generally,
the scaled matrix, $\sqrt{ \rho_i/\rho_j } \, w_{ij}$,
to be symmetric.
%
This symmetry allows the diagonalization
of $\mathbf w$ with a set of
eigenvectors, $\phi_{ki}$,
%
\begin{equation}
  \sum_{i = 1}^n \phi_{ki} \, w_{ij}
  =
  \lambda_k \, \phi_{kj}
  ,
\label{eq:eig_w}
\end{equation}
%
satisfying the orthonormal conditions\cite{vankampen}:
%
\begin{align}
  \sum_{k = 0}^{n - 1}
    \phi_{ki} \, \phi_{kj}
  &=
  \delta_{ij} \, \rho_i,
  \label{eq:eig_orthonormal_cols}
  \\
  \sum_{i = 1}^n
    \frac{ \phi_{ki} \, \phi_{li} }
         { \rho_i }
  &=
  \delta_{kl}
  ,
  \label{eq:eig_orthonormal_rows}
\end{align}
%
where we have set the index of the first eigenmode to $0$
instead of $1$ for later convenience.

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \label{eq:w_balance}
\end{equation}
%
We will label this eigenvector as the first one,
and
%
\begin{equation}
  \phi_{0i} = \rho_i,
\label{eq:eigenmode0}
\end{equation}
%
with $\lambda_0 = 1$.
%
Clearly, Eq. \eqref{eq:eigenmode0}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 0$,
while the orthogonality demands
%
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \delta_{k0}
  .
\label{eq:ortho1}
\end{equation}
%
We may write the updating matrix in $\phi_{ki}$'s
as\cite{bussi2006}
%
\begin{equation}
  w_{ij}
  =
  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
  \lambda_k \, \phi_{ki} \, \phi_{kj}
  .
  \notag
\end{equation}

@@

We can further define
a generalized Fourier transform, $\mathcal{F}$,
for variable, $A_i$,
%
\begin{equation}
  {\tilde A}_k
  \equiv \mathcal{F}[A_i]_k
  \equiv \sum_{i = 1}^n \phi_{ki} \, A_i
  ,
  \label{eq:gft_def}
\end{equation}
%
and the inverse transform is
%
\begin{equation}
  A_i =
  \frac{1}{\rho_i} \sum_{k = 1}^n \phi_{ki} \, \tilde{A}_k
  .
  \label{eq:gft_inv}
\end{equation}
%
Using this definition,
we can rewrite Eq. \eqref{eq:ortho1} as
%
\begin{equation}
  \mathcal F[1]_k = \delta_{k0}
  ,
  \label{eq:ortho0}
\end{equation}
%
and the bin average, Eq. \eqref{eq:vav_def} can be written as
\begin{equation}
  v_\mathrm{av} = \mathcal F[v_i]_0 = \tilde v_0
  .
  \label{eq:vav_vt0}
\end{equation}

We can now diagonalize
the multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
as
%
$$
{\tilde v}_k(t + 1) =
{\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
{\tilde \eta}_k(t)
,
$$
or in the continuous-time form,
%
\begin{equation}
  \dot{\tilde v}_k(t)
  =
  \alpha(t) \, \lambda_k \, {\tilde \eta}_k(t)
  .
  \label{eq:ut_diffeq_mbin}
\end{equation}
%
Deducting the bin average, we get
%
$$
\begin{aligned}
  \dot {\tilde x}_k(t)
  &= \alpha(t) \, \lambda_k
  \left[
    {\tilde \eta}_k(t)
    -
    {\tilde \eta}_\mathrm{av}(t) \, \delta_{k0}
  \right]
  \\
  &\approx
  -\alpha(t) \, \lambda_k \,
  \bigl[ {\tilde x}_k(t) - {\tilde \xi}_k (t) \bigr]
  ,
\end{aligned}
$$
where we have used the linear approximation
introduced in Sec. \ref{sec:sbin_diffeq}.

@@

The error defined in Eq. \eqref{eq:error_def}
can be rewritten in terms of the eigenmodes of $\mathbf w$,
%
\begin{align}
  \Err
  \equiv
  \sum_{i = 1}^n \rho_i \,
    \left\langle x_i^2 \right\rangle
  =
  \sum_{i, j = 1}^n \sum_{k=0}^{n-1}
   \phi_{ki} \, \phi_{kj} \,
    \left\langle x_i \, x_j \right\rangle
  =
  \sum_{k = 0}^{n - 1}
    \left\langle
      {\tilde x}_k^2
    \right\rangle
  ,
\notag
\end{align}
%
where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
%
From Eq. \eqref{eq:eigenmode0},
the first eigenmode,
${\tilde x}_0 = \sum_{j=1}^n \rho_j \, x_j$,
vanishes identically by Eq. \eqref{eq:rhox_sum}.
%
Thus, we can start the last sum %in Eq. \eqref{eq:xt2_sum}
from $k = 1$,
%
\begin{equation}
  \Err(T)
  =
  \sum_{k = 1}^{n - 1}
    \left\langle
      {\tilde x}_k^2(T)
    \right\rangle
  .
\label{eq:xt2_sum}
\end{equation}
